Using Probabilistic Knockoffs of Binary Variables to Control the False Discovery Rate

### 2015-07-30 masters_paper/paper.tex
Done with discussion
Updating section 1

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
    Variable selection is an essential problem to fitting a regression model. For linear regression, the knockoff filter (Barber and Cand\`{e}s, 2014) offers a method of exact FDR control. However, the method as originally formulated does not extend to other generalized linear models such as logistic regression. This paper offers an extension of this method which will work with binary features in the context of linear regression, but also will extend to other GLMs.

\subsection{Knockoff Filter}
    Let us consider the consider the usual setting for linear regression, where $n$ observations of a variable of interest $y$ arise from the model
    \[\mathbf{y} = X\beta + \mathbf{z}\]
    where $\mathbf{y}\in\R^n$ are the observed values of $y$, $X \in \R^{n\times p}$ is the matrix of predictor variables, $\beta\in\R^p$ are the unknown coefficients, and $\mathbf z$ is Gaussian noise. It is important to note that a intercept vector $\mathbf 1$ is not included in $X$. For the purpose of this paper, only the case where $2p\leq n$ will be considered, though the methods can be extended to $p\leq n$. I will to refer the $j$th column of $X$ as $X_j$, which is $n$ observations of the random variable $x_j$. The random variable $x_j$ is in turn $j$th entry of the random vector valued variable $\mathbf x$. In the situation where $\beta$ is sparse and there is reason to expect that $x_j$ and $y$ are uncorrelated for many $j$, the knockoff filter is a method to select a subset of the $x_j$ which are likely correlated with $y$. \par
    The method specifically seeks to control the false discovery rate (FDR), which is the expected portion of some selection of $x_j$ which are, in fact, uncorrelated with $y$. For a procedure which chooses a subset of variables $\hat S\subseteq \{1,\ldots,p\}$, this is defined as  
    \[ \textrm{FDR} = \E\left[\frac{\vert{\{j: \beta_j=0 \; \& \; j\in\hat S\}}\vert}{\max\{\vert{\hat S}\vert,1\}} \right] \]
    FDR is controlled at level $q$ if FDR is less than $q$ irrespective of the coefficients $\beta$. \par

    The knockoff filter achieves FDR control in several steps, the first of which is creating a set of `knockoff' features $\tilde X_j$ which imitate the original $X_j$ while being less correlated with $y$. In particular, the matrix of knockoffs $\tilde X$ has the same internal correlation structure as $X$, the knockoff feature $\tilde X_j$ has the same correlation with other $X_i$ as $X_j$ does, but the correlation between $\tilde X_j$ and $X_j$ is minimized. In other words, where $ G := X^TX$, 

@@

    for some vector $s\in \R^p$ such that, where $\diag{G}$ is the vector made from the main diagonal of $G$, $\diag{G} - \mathbf s$ is small. Since $\tilde X_j$ and $X_j$ have relatively low correlation, as long as $X_j$ is created independently of $\mathbf y$, $\tilde X_j$ will have lower correlation with $\mathbf y$ than $X_j$. In the case where $X_j$ is uncorrelated with $\mathbf y$, $\tilde X_j$ will also be uncorrelated with $\mathbf y$. How these knockoffs are actually generated will be described shortly.\par
    The second step is to fit a series of Lasso model of $\mathbf y$ on the combined design matrix $X_L=[X \; \tilde X]$ so as to determine the largest value $\lambda$ at which the coefficient for each $X_j$ and $\tilde X_j$ is nonzero. The subscript $L$ refers to the large design matrix with both features and knockoffs. \footnote{Other statistics besides the $\lambda$s can also be used in the knockoff method, but are not considered in this paper} Recall, for a given $\lambda$, the estimated coefficient from the Lasso regression will be 
    \[ \beta(\lambda) = \argmin_\mathbf b \left\{\frac{1}{2}\|\mathbf y - X_L\mathbf b\|^2_2 + \lambda\|b\|_1 \right\}\]
    In the general case, the coefficient for one feature in a Lasso regression being first nonzero for a larger value of $\lambda$ than another feature is an indication that the first feature is a stronger predictor of the outcome than the second. Thus, since, by construction, the knockoff features are weaker predictors than the originals, we would expect the original features to have nonzero coefficients sooner than the knockoff features when the original is a valid predictor. On the other hand, since $\beta(\lambda)$ only depends on $X_L$ through the sufficient statistics $G_L=X_L^TX_L$ and $X_L^T\mathbf y$, for null predictors, the coefficients of the original features won't enter any sooner on average than knockoff feature. This can be seen by switching a null $X_j$ with $\tilde X_j$; by construction, $G_L$ will be unaltered, while $\E\left[X_L^T\mathbf y\right]$ will also be unchanged, since $\E[X_j^T\mathbf y]=\E[\tilde X_j^T\mathbf y]=0$. \par

@@

\subsection{Paper Organization}
The rest of this paper will be organized in the following fashion:
\begin{itemize}
    \item Section 2 will go into more detail about how the original knockoffs breakdown with generalized linear models. In particular, several simulations will demonstrate how they breakdown.
    \item Section 3 will develop the method and theory for generating binary knockoffs.
    \item Section 4 will discuss tests of Binary knockoffs in simulation, both in comparison to the original knockoffs with Lasso and for logistic regression by itself.
    \item Section 5 is the final section, and will contain discussion of the results as well as areas for further work.
\end{itemize}

\section{Issues With Deterministic Knockoffs} 
    The knockoff filter and $W$ statistics have a very natural extension to generalized linear models. Here, where $l(\beta\st X, \tilde X)$ is the log likelihood of coefficients $\beta$ given the model, the $L1$ regularized regression model will have estimated coefficient vector
    \[ \beta(\lambda) = \argmin_\mathbf b \left\{l(\beta\st X, \tilde X) + \lambda\|b\|_1 \right\}\]
    The $Z_j$ and $\tilde Z_j$ are then, once again, the largest $\lambda$ such that the given original or knockoff feature has a positive coefficient, and then $W_j$ is constructed exactly as before. The problem with this is that if the given GLM doesn't have the same sufficient statistics as linear regression, there is no theoretical guarantee that the knockoff filter will control FDR. This is since, for null $X_j$, there is no theory to guarantee the $Z_j$ and $\tilde Z_j$ are exchangeable, and thus $W_j$ is not guaranteed to be symmetrically distributed around $0$. \par
    \begin{figure}[h]
        \begin{center}
        \includegraphics[width=14cm]{images/entryrate_original_logit}
    \end{center}
        \caption{The mean cumulative portion of $W$ statistic which are positive by absolute rank of the value. The mean is over 1000 simulations. On the left side, the features are drawn from a random standard normal distribution. On the right side, they are drawn as Bernoulli with probability $.25$ of being 1. The top has 50 features uncorrelated with the outcome, while the bottom has 40 features uncorrelated with the outcome and 10 correlated with it.}
    \end{figure}

@@

    However, this method, with the real possibility of extending the knockoff filter to GLMs, offers an interesting development if these issues can be resolved. For the issue with the deviation from the desired cross-moment matrix, a solution would likely take the form of an ensemble method. Rather than fit the $A$ and $P$ matrix to generate one $\tilde X$, several $\tilde X$ might be generated and used to make variable selections $S_m$. A finally variable selection $S$ could be made based on theses, perhaps only selecting the variables selected by the majority of the $S_m$. This, or some other sort of average across multiple realizations of $\tilde X$, could likely be used to dampen the effect of variance in the generation of $\tilde X$. This would at least make binary knockoffs reliable for Lasso. \par

@@

    The final area of concern is how binary knockoffs will occur when the $X$ matrix is not fit well by the Ising model due to significant high order interactions. Hopefully, having matched the first and second moments of $X$, binary knockoffs will continue to provide reasonable FDR control. An avenue to deal with this though would be to explicitly model these higher order interactions. One might introduce higher order moments into the fitting of $\mu$-conditionals. For instance, if $A$ is a three dimensional matrix whose entry $a_{ijk}$ would be added to the linear predictor when $x_i=x_j=x_k=1$. Just as the $\mu$-conditionals are based on a regression of $X_i$ on $X_{1:i-1}$ with no interactions, this would be akin to adding two way interactions. This can be used to replicate arbitrary binary vectors to an arbitrary desired level of precision. However, in practice this has big limitations. For one, adding additional levels of interactions will make the computation much more difficult very quickly. As well, as far as generating knockoffs, it will become more and more difficult to reduce the correlation between $x_i$ and $\tilde x_i$ as higher order interactions are insisted on. Nonetheless, this may be an avenue for further investigation.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Variable selection is an essential problem to fitting a regression model. For linear regression, the knockoff filter \cite{2014arXiv1404.5609F} offers a method of exact FDR control. However, the method as originally formulated does not extend to other generalized linear models (GLMs) such as logistic regression. This paper offers an extension of this method for binary features in the context of both GLMs as well as linear regression.

\subsection{Background: the Knockoff Filter}
    Let us consider the usual setting for linear regression, where $n$ observations of a variable of interest $y$ arise from the model
    \[\mathbf{y} = X\beta + \mathbf{z}\]
    where $\mathbf{y}\in\R^n$ are the observed values of $y$, $X \in \R^{n\times p}$ is the matrix of predictor variables, $\beta\in\R^p$ are the unknown coefficients, and $\mathbf z$ is Gaussian noise.It is important to note that a intercept vector $\mathbf 1$ is not included in $X$. For the purpose of this paper, only the case where $2p\leq n$ will be considered, though the methods can be extended to $p\leq n$. I will to refer the $j$th column of $X$ as $X_j$, which is $n$ observations of the random variable $x_j$. The random variable $x_j$ is in turn $j$th entry of the random vector valued variable $\mathbf x$. Let $\beta$ be sparse, implying only a subset of the features $x_j$ have an effect on  $y$. In other words, there is a true model $S\subseteq \{1,\ldots,p\}$ such that $\beta_i\neq0$ if and only if $i\in S$. \par
     In this context, the knockoff filter is designed to choose a model $\hat S$ so as to control the false discovery rate (FDR), or the portion of variables chosen by $\hat S$ which aren't in the true model $S$:
    \[ \textrm{FDR} = \E\left[\frac{\vert{\{j: j\not\in S \; \& \; j\in\hat S\}}\vert}{\max\{\vert{\hat S}\vert,1\}} \right] \]
    In other words, FDR is the portion of features which are thought to have an effect on $y$ but actually have no effect. FDR is controlled at level $q$ if FDR is less than $q$ irrespective of the coefficients $\beta$. \par

    The knockoff filter achieves FDR control in several steps, the first of which is creating a set of `knockoff' features $\tilde X_j$ which imitate the original $X_j$ while being no more correlated with $y$. In particular, the matrix of knockoffs $\tilde X$ has the same internal correlation structure as $X$ and the knockoff feature $\tilde X_j$ has the same correlation with other $X_i$ as $X_j$ does, but the correlation between $\tilde X_j$ and $X_j$ is minimized. In other words, for $G := X^TX$, 

@@

    for some vector $\mathbf{s}\in \R^p$ such that, writing $\diag{G}$ as the vector made from the main diagonal of $G$, $\diag{G} - \mathbf s$ is small. Since $\tilde X_j$ and $X_j$ have relatively low correlation, as long as $X_j$ is created independently of $\mathbf y$, $\tilde X_j$ will have lower correlation with $\mathbf y$ than $X_j$. In the case where $X_j$ is a null predictor of $\mathbf y$, $\tilde X_j$ will thus also be a null predictor of $\mathbf y$. How these knockoffs are actually generated will be described shortly.\par
    The second step is to fit a series of LASSO models of $\mathbf y$ on the combined design matrix $X_L:=[X \; \tilde X]$ so as to determine the largest value $\lambda$ at which the coefficient for each of $X_j$ and $\tilde X_j$ is nonzero.\footnote{Other statistics besides the $\lambda$s can also be used in the knockoff method, but are not considered in this paper} The subscript $L$ refers to the large design matrix with both features and knockoffs. Recall, for a given $\lambda$, the estimated coefficient from the LASSO regression will be 
    \[ \beta(\lambda) = \argmin_\mathbf b \left\{\frac{1}{2}\|\mathbf y - X_L\mathbf b\|^2_2 + \lambda\|\mathbf{b}\|_1 \right\}\]
    In the general case, the ``LASSO path" refers to the path of solutions as $\lambda$ is decreased from $\infty$ to $0$, and we say that a features ``enters the path" at $\lambda$ when its coefficient is first estimated to be nonzero. One feature entering the LASSO path before another is an indication that the first feature is a stronger predictor of the outcome than the second. Thus, since, by construction, the knockoff features are weaker predictors than the originals, we would expect the original features to enter the LASSO path sooner than the knockoff features when the original is a valid predictor. On the other hand, since $\beta(\lambda)$ only depends on $X_L$ through the sufficient statistics $G_L:=X_L^TX_L$ and $X_L^T\mathbf y$, for null predictors, the coefficients of the original features won't enter any sooner on average than knockoff feature. This can be seen by switching a null $X_j$ with $\tilde X_j$; by construction, $G_L$ will be unaltered, while $\E\left[X_L^T\mathbf y\right]$ will also be unchanged, since $\E[X_j^T\mathbf y]=\E[\tilde X_j^T\mathbf y]=0$. \par

@@

\subsection{Paper Organization}
The rest of this paper will be organized in the following fashion:
\begin{itemize}
    \item Section 2 will go into more detail about how the original knockoffs breakdown with generalized linear models. In particular, several simulations will demonstrate how they breakdown.
    \item Section 3 will develop the method and theory for generating binary knockoffs.
    \item Section 4 will discuss tests of Binary knockoffs in simulation, both in comparison to the original knockoffs with LASSO and for logistic regression by itself.
    \item Section 5 is the final section, and will contain discussion of the results as well as areas for further work.
\end{itemize}

\section{Issues With Deterministic Knockoffs} 
    The knockoff filter and $W$ statistics have a very natural extension to generalized linear models. Here, where $l(\beta\st X, \tilde X)$ is the log likelihood of coefficients $\beta$ given the model, the $L1$ regularized regression model will have estimated coefficient vector
    \[ \beta(\lambda) = \argmin_\mathbf b \left\{l(\beta\st X, \tilde X) + \lambda\|b\|_1 \right\}\]
    The $Z_j$ and $\tilde Z_j$ are then, once again, the largest $\lambda$ such that the given original or knockoff feature has a positive coefficient, and then $W_j$ is constructed exactly as before. The problem with this is that if the given GLM doesn't have the same sufficient statistics as linear regression, there is no theoretical guarantee that the knockoff filter will control FDR. This is since, for null $X_j$, there is no theory to guarantee the $Z_j$ and $\tilde Z_j$ are exchangeable, and thus $W_j$ is not guaranteed to be symmetrically distributed around $0$. \par
    \begin{figure}[h]
        \begin{center}
        \includegraphics[width=14cm]{images/entryrate_original_logit}
    \end{center}
        \caption{The mean cumulative portion of $W$ statistic which are positive by absolute rank of the value. The mean is over 1000 simulations. On the left side, the features are drawn from a random standard normal distribution. On the right side, they are drawn as Bernoulli with probability $.25$ of being 1. The top has 50 null features, while the bottom has 40 null features with the outcome and 10 correlated with it.}
    \end{figure}

@@

    However, this method, with the real possibility of extending the knockoff filter to GLMs, offers an interesting development if these issues can be resolved. For the issue with the deviation from the desired cross-moment matrix, a solution would likely take the form of an ensemble method. Rather than fit the $A$ and $P$ matrix to generate one $\tilde X$, several $\tilde X$ might be generated and used to make variable selections $S_m$. A finally variable selection $S$ could be made based on theses, perhaps only selecting the variables selected by the majority of the $S_m$. This, or some other sort of average across multiple realizations of $\tilde X$, could likely be used to dampen the effect of variance in the generation of $\tilde X$. This would at least make binary knockoffs reliable for LASSO. \par

@@

    The final area of concern is how binary knockoffs will occur when the $X$ matrix is not fit well by the Ising model due to significant high order interactions. Hopefully, having matched the first and second moments of $X$, binary knockoffs will continue to provide reasonable FDR control. An avenue to deal with this though would be to explicitly model these higher order interactions. One might introduce higher order moments into the fitting of $\mu$-conditionals. For instance, if $A$ is a three dimensional matrix whose entry $a_{ijk}$ would be added to the linear predictor when $x_i=x_j=x_k=1$. Just as the $\mu$-conditionals are based on a regression of $X_i$ on $X_{1:i-1}$ with no interactions, this would be akin to adding two way interactions. This can be used to replicate arbitrary binary vectors to an arbitrary desired level of precision. However, in practice this has big limitations. For one, adding additional levels of interactions will make the computation much more difficult very quickly. As well, as far as generating knockoffs, it will become more and more difficult to reduce the correlation between $x_i$ and $\tilde x_i$ as higher order interactions are insisted on. Nonetheless, this may be an avenue for further investigation. \par
